---
title: "posterior_analysis"
output: html_document
---

```{r}
rm(list=ls())
library(ggplot2)
library(dplyr)
library(coda)
library(aida)
```

```{r}
theme_set(theme_aida())

##################################################
## CSP-colors
##################################################
CSP_colors = c(
  "#7581B3", "#99C2C2", "#C65353", "#E2BA78", "#5C7457", "#575463",
  "#B0B7D4", "#66A3A3", "#DB9494", "#D49735", "#9BB096", "#D4D3D9",
  "#414C76", "#993333", "#A0A0A0"
  )
# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = CSP_colors)
}
scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = CSP_colors)
}
```

Helper functions.
```{r}
entropy <- function(p) {
  p <- p[p > 0]  # Avoid log(0)
  -sum(p * log2(p))
}

compute_log_lik <- function(true, predicted, epsilon = 1e-10) {
  ifelse(true == predicted, log(1), log(epsilon))
}
```

# Import data
```{r}
posteriorSamples_gb <- read.csv("production_posterior_full_gb_10k_4p.csv")
posteriorSamples_inc <- read.csv("production_posterior_full_inc_10k_4p.csv")

posteriorPredictive_gb <- read.csv("production_posteriorPredictive_full_gb_10k_4p.csv")
posteriorPredictive_gb_svi_soft <- read.csv("production_posterior_gb_relevanceXsharpness_SVI_long_soft.csv")
posteriorPredictive_inc_svi_soft <- read.csv("production_posterior_inc_relevanceXsharpness_SVI_long_soft.csv")
posteriorPredictive_inc <- read.csv("production_posteriorPredictive_full_inc_10k_4p.csv")
```

```{r}
posteriorPredictive_gb_svi_soft$relevant_property <- as.factor(posteriorPredictive_gb_svi_soft$relevant_property)
posteriorPredictive_gb_svi_soft$annotation_category <- as.factor(posteriorPredictive_gb_svi_soft$annotation_category)
posteriorPredictive_gb_svi_soft$sharpness <- as.factor(posteriorPredictive_gb_svi_soft$sharpness)

posteriorPredictive_inc_svi_soft$relevant_property <- as.factor(posteriorPredictive_inc_svi_soft$relevant_property)
posteriorPredictive_inc_svi_soft$annotation_category <- as.factor(posteriorPredictive_inc_svi_soft$annotation_category)
posteriorPredictive_inc_svi_soft$sharpness <- as.factor(posteriorPredictive_inc_svi_soft$sharpness)

posteriorPredictive_gb_svi_soft %>%
ggplot(aes(x = annotation_category, y = probability, fill = source)) +
  geom_bar(stat= "identity", position = "dodge") +
  facet_wrap(sharpness ~ relevant_property, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

posteriorPredictive_inc_svi_soft %>% filter(source == "model") %>%
ggplot(aes(x = annotation_category, y = probability, fill = source)) +
  geom_bar(stat= "identity", position = "dodge") +
  #geom_errorbar(aes(ymin = mean_probability - sd_probability, ymax = mean_probability + sd_probability)) +
  facet_wrap(sharpness ~ relevant_property, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(posteriorPredictive_inc_svi_soft, aes(x = probability, y = source)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(sharpness ~ relevant_property, scales = "free_y")
```



# Old stuff
Convert posterior samples to MCMC object
```{r}
posteriorSamples_gb_mcmc <- as.mcmc(posteriorSamples_gb)
posteriorSamples_inc_mcmc <- as.mcmc(posteriorSamples_inc)
```

# Posterior analysis

## Global Speaker
```{r}
plot(posteriorSamples_gb)
summary(posteriorSamples_gb_mcmc)
traceplot(posteriorSamples_gb_mcmc)
autocorr.plot(posteriorSamples_gb_mcmc)
effectiveSize(posteriorSamples_gb_mcmc)
```
Density Plot of Productions with empirical data, facet by conditions, they should normalise to 1 and stack up together on y-axis
```{r}
library(ggridges)

posteriorPredictive_gb %>%
  count(conditions, annotation, sharpness) %>%
  group_by(conditions) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(aes(x = annotation, y = prop, fill = annotation)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(conditions~sharpness) +
  labs(title = "Normalized Production Distributions by Condition",
       y = "Proportion", x = "Production (Annotation)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
library(ggridges)
library(forcats)

# Step 1: Prepare and normalize
summary_empirical <- posteriorPredictive_gb %>%
  count(conditions, annotation, sharpness) %>%
  group_by(conditions) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Step 2: Reorder annotation within each condition by prop
summary_empirical <- summary_empirical %>%
  group_by(conditions, sharpness) %>%
  mutate(annotation_reordered = fct_reorder(annotation, prop, .desc = TRUE)) %>%
  ungroup()

# Step 3: Plot
ggplot(summary_empirical, aes(x = prop, y = annotation_reordered, fill = annotation)) +
  #geom_density_ridges(stat = "identity", scale = 0.9, height = 10 * prop) +
  geom_col() + 
  facet_wrap(conditions~sharpness, scales = "free_y") +
  labs(
    title = "Normalized Production Distributions by Condition",
    x = "Proportion of Posterior Samples",
    y = "Production Category (ordered by frequency)"
  ) +
  scale_fill_manual(values = rev(CSP_colors)) +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 8)
  )

# Step 4: compute entropy
empirical_entropy_by_condition <- summary_empirical %>%
  group_by(conditions, sharpness) %>%
  summarise(entropy = entropy(prop), .groups = "drop")
```

Unlist the predictions and convert them to strings.

```{r}
library(tidyr)
library(purrr)

# String-to-index mapping (from your previous conversion)
converted_dict <- c(
  "D" = 0, "DC" = 1, "DCF" = 2, "DF" = 3, "DFC" = 4,
  "C" = 5, "CD" = 6, "CDF" = 7, "CF" = 8, "CFD" = 9,
  "F" = 10, "FD" = 11, "FDC" = 12, "FC" = 13, "FCD" = 14
)

# Reverse it: index (as character) → string
index_to_string <- setNames(names(converted_dict), as.character(converted_dict))

# Unnest predictions (creates 2101 * 6000 rows)
pred_wide <- posteriorPredictive_gb %>%
  select(sharpness, conditions, predictions, annotation) %>%
  # Get rid of "[" and "]"
  mutate(
  predictions = gsub("\\[|\\]", "", predictions),      # remove [ and ]
  predictions = gsub("^\\s+|\\s+$", "", predictions), # remove leading/trailing whitespace
  predictions = strsplit(predictions, ",\\s*"),
  predictions = map(predictions, as.numeric)
) %>%
  unnest_wider(predictions, names_sep = "_draw") %>%
  mutate(across(-c(sharpness, conditions), ~ index_to_string[as.character(.)]))

# Add posteriorPredictive_gb$annotation as a col to pred_wide, called empirical
pred_wide <- pred_wide %>%
  mutate(empirical = posteriorPredictive_gb$annotation) %>%
  select(sharpness, conditions, starts_with("predictions"), empirical)
```

Compute the probs of categorical distribution per draw and per sharpness. 

```{r}
# 1. Pivot prediction draws to long format
pred_long <- pred_wide %>%
  pivot_longer(
    cols = starts_with("predictions"),
    names_to = "draw",
    values_to = "prediction"
  )

# 2. Count frequency of each predicted category per condition x annotation x draw
pred_probs <- pred_long %>%
  group_by(sharpness, conditions, draw, prediction) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(sharpness, conditions, draw) %>%
  mutate(prob = n / sum(n)) %>%
  ungroup()

# 3. Ensure all 15 categories are present (fill missing ones with 0)
all_categories <- sort(unique(pred_long$prediction))  # or supply manually

pred_probs_complete <- pred_probs %>%
  complete(
    sharpness,
    conditions,
    draw,
    prediction = all_categories,
    fill = list(prob = 0)
  )
```

Aggragate the probs by sharpness and draw.
```{r}
se <- function(x) sd(x) / sqrt(length(x))

pred_probs_agg <- pred_probs_complete %>%
  group_by(sharpness, conditions, prediction) %>%
  summarise(mean_prob = mean(prob), se_prob = se(prob), .groups = "drop") %>%
  arrange(sharpness, prediction)

ggplot(pred_probs_agg, aes(x = mean_prob, y = prediction, fill = prediction)) +
  #geom_density_ridges(stat = "identity", scale = 0.9, height = 10 * prop) +
  geom_col() + 
  # add error bars
  geom_errorbarh(aes(xmin = mean_prob - se_prob, xmax = mean_prob + se_prob), width = 0.02) +
  facet_wrap(conditions~sharpness, scales = "free_y") +
  labs(
    title = "Normalized Production Distributions by Condition",
    x = "Proportion of Posterior Samples",
    y = "Production Category (ordered by frequency)"
  ) +
  scale_fill_manual(values = rev(CSP_colors)) +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 8)
  )
```
Compute Entropy of the distribution given each draw, conditions, sharpness.
```{r}


entropy_by_draw <- pred_probs_complete %>%
  group_by(sharpness, conditions, draw) %>%
  summarise(entropy = entropy(prob), .groups = "drop")

entropy_by_draw %>% group_by(sharpness, conditions) %>%
  # Compute mean and standard deviation of entropy for each condition and sharpness)
summarise(
  mean_entropy = mean(entropy),
  sd_entropy = sd(entropy),
  .groups = "drop"
) -> entropy_predictions_summary

entropy_predictions_summary %>% mutate(empirical_entropy = empirical_entropy_by_condition$entropy) -> entropy_predictions_summary

ggplot(entropy_predictions_summary, aes(x = sharpness, y = mean_entropy, color = conditions)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = mean_entropy - sd_entropy, ymax = mean_entropy + sd_entropy), width = 0.2) +
  labs(
    title = "Entropy of Predictions",
    x = "Sharpness",
    y = "Mean Entropy"
  ) +
  geom_point(aes(y = empirical_entropy), shape = 4, size = 3) +
  scale_color_manual(values = CSP_colors)
```

Compute LOO.
```{r}
library(loo)
# Step 1: Compute matrix of log-lik per sample
log_lik_matrix <- pred_wide %>%
  select(starts_with("predictions")) %>%
  as.data.frame() %>%
  map_df(~ compute_log_lik(pred_wide$empirical, .x)) %>%
  as.matrix() %>%
  t()  # transpose to: draws x data points

loo_result <- loo(log_lik_matrix)
print(loo_result)
```


## Incremental Speaker
```{r}
plot(posteriorSamples_inc)
summary(posteriorSamples_inc_mcmc)
traceplot(posteriorSamples_inc_mcmc)
autocorr.plot(posteriorSamples_inc_mcmc)
effectiveSize(posteriorSamples_inc_mcmc)
```
Look into the correlation between parameters
```{r}
cor_matrix <- cor(posteriorSamples_inc)
print(cor_matrix)
```


```{r}
x <- data$predictions[1]

# X is a vector of predictions stored in strings, we need to convert it to a list of numbers first
# first, get rid of [ at the beginning and ] at the end
convert_to_list <- function(x) {
  x <- gsub("\\[|\\]", "", x)
  # then split the string by comma
  x <- strsplit(x, ",")[[1]]
  # convert to numeric
  x <- as.numeric(x)
  return(x)
}

# apply to every row in data$predictions and store the results in a new dataframe called predictions
data$predictions %>% sapply(convert_to_list)
# with row now the values, and column the index

```

```{r}
# Extract unique statesArray in data and corresponding annotation_seq_flat
data %>% group_by(annotation) %>% ggplot(aes(x = annotation, fill = conditions)) +
  geom_bar()

data %>% group_by(annotation) %>% ggplot(aes(x = predictions, fill = conditions)) +
  geom_density()
```

```{r}
# Compute by condition mean
data %>%
  group_by(conditions) %>%
  summarise(mean_annotation = mean(annotation_seq_flat),
            mean_predictions = mean(mean_predictions)) -> by_condition_mean

ggplot(by_condition_mean, aes(x = mean_annotation, y = mean_predictions)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()
```

