---
title: "analysis"
output: html_document
---

Supress messages in this document so that the output is cleaner.
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```
Import packages.
```{r}
rm(list = ls())
options(warn = -1)
library(tidyverse)
library(lme4)
library(lmerTest)
library(ggplot2)
library(aida)
library(BayesFactor)
library(pwr)
library(brms)
library(emmeans)
library(effsize)
library(HDInterval)
```

Set up theme for ggplot.

```{r}
theme_set(theme_aida())

##################################################
## CSP-colors
##################################################
CSP_colors = c(
  "#7581B3", "#99C2C2", "#C65353", "#E2BA78", "#5C7457", "#575463",
  "#B0B7D4", "#66A3A3", "#DB9494", "#D49735", "#9BB096", "#D4D3D9",
  "#414C76", "#993333"
  )
# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = CSP_colors)
}
scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = CSP_colors)
}
```
Set up plots dir.
```{r}
plots_dir <- "./plots"
```

# Data Import & Preprocessing
Import data.
```{r}
data <- read.csv("../../01-dataset/00-slider-data-replication-raw.csv")
```

Show basic information.
```{r}
#show how many pps
sum(xtabs(~id,data=data)/180)

# show how many lists
round(xtabs(~list, data=data)/180)

```
Set up factors.
```{r}
data$conditions <- droplevels(as.factor(data$conditions))
data$id <- droplevels(as.factor(data$id))
data$slider_value <- as.numeric(data$slider_value)
data$dist <- ifelse(data$list<4, "sharp", "blurred")
data$dist <- droplevels(as.factor(data$dist))
```
Deal with missing data 
```{r}
#coercion slider value when missing list to NA (due to unexpected technical errors).
missing_list <- which(is.na(data$list)==TRUE)
data[missing_list,]$slider_value <- NA

#coercion slider value when none fit to NA. This refers to the third option "neither".
data$slider_value <- ifelse(data$slider_value==-1, NA, data$slider_value)

```
Subset filler from main data set.
```{r}
data_filler <- subset(data, str_sub(data$conditions, 1, 1) == "f")
data <- droplevels(subset(data, str_sub(data$conditions, 1, 1) != "f"))
```
Encode slider. The max (100) position is randomized on the slider, either Left or Right. For Left, we have 1 -100 to get the same result as Right, which is the default. 
```{r}
data$prefer_first_1st <- ifelse(data$leftright_trial=="1left", 100-data$slider_value, data$slider_value)
data$prefer_first_1st <- data$prefer_first_1st - 50 # We further extract 50 to get the range from -50 to 50 for better visualization.
```
Encode condition's name into English.
```{r}
data$combination <- as.factor(ifelse(str_sub(data$conditions, 3, 4)=="cf", "color_form",
                           ifelse(str_sub(data$conditions, 3, 4)=="dc", "dimension_color",
                                  "dimension_form")))


data$relevant_property <- as.factor(ifelse(str_sub(data$conditions, 1, 1)== "e", "first",
                                           ifelse(str_sub(data$conditions, 1, 1)== "z", "second",
                                           "both")))
```


# Descriptive Statistics
Aggregate data.
```{r}
#aggregate data
aggregate(data$prefer_first_1st, list(data$conditions), FUN = function(x){c(mean(x), sd(x)/sqrt(length(x)))})
aggregate(data$prefer_first_1st, list(data$conditions, data$dist), FUN = function(x){c(mean(x), sd(x)/sqrt(length(x)))})

ratings_aggregated<-aggregate(prefer_first_1st~combination+relevant_property+dist, data=data, mean)
#TODO: find another way to compute ses (SEwithin?)
ses_ratings<-aggregate(prefer_first_1st~combination+relevant_property+dist, data=data, function(x) {sd(x)/sqrt(length(x))})
ratings_aggregated<-cbind(ratings_aggregated,ses_ratings[,4])

names(ratings_aggregated)[5]<-"se"

```
Draw a distribution plot of slider.
```{r}
# Raw data
ggplot(data, aes(x = prefer_first_1st, fill = relevant_property)) +
  geom_density(fill = CSP_colors[1]) 

```

Draw a summary plot of slider by conditions.
```{r}
# make bar plot for descriptive analysis
p <- ggplot(ratings_aggregated, aes(x=combination, y=prefer_first_1st, fill=relevant_property)) +
  ylab("preference for the position in front")+
  geom_bar(stat="identity", position=position_dodge()) +
  geom_errorbar(aes(ymin=prefer_first_1st-se, ymax=prefer_first_1st+se), width=.2,
                position=position_dodge(.9))+
  facet_wrap(~dist)+
  scale_fill_brewer(palette="Paired") + 
  theme(axis.text.x = element_text(angle = 45, hjust=1))
p
ggsave(paste0(plots_dir, "/descriptive_plot.png"), plot = p, device = "png", width = 10, height = 10, units = "in", dpi = 300)

```
# Inferential Statistics
Normality test.
```{r}
nortest::ad.test(data$prefer_first_1st) # Not passed with raw data. -> We need to transform the data.
```
Data transformation.
```{r}
data$sli<-(data$prefer_first_1st+50.5)/101
model_raw <- lmer(data$prefer_first_1st~relevant_property*combination*dist
           +(1|id)
           +(1|item), data=data)
model_log <- lmer(log(sli)~relevant_property*combination*dist
           +(1|id)
           +(1|item), data=data)
data$transli<-qlogis(jitter(data$sli))
model_logit <- lmer(transli~relevant_property*combination*dist
             +(1|id)
             +(1|item), data=data)
par(mfrow=c(1,3))
qqnorm(residuals(model_raw), main = "default value")
qqnorm(residuals(model_log), main = "value log transformed")
qqnorm(residuals(model_logit), main = "value logit transformed")

# We will use logit transformed data for further analysis.
model_full <- model_logit
```
```{r}
step(model_full) # No sig. three way interaction. We can reduce model
model_full <- update(model_full, .~.-relevant_property:combination:dist)
```
We want to resolve the interaction between distribution and combination first.
```{r}
# Subset the data to dimension_X and color_X
data_dimisionX <- subset(data, combination=="dimension_form" | combination=="dimension_color")
data_colorForm <- subset(data, combination=="color_form")

# Fit two models for each subset
model_dimisionX <- lmer(transli~relevant_property*dist
             +(1|id)
             +(1|item), data=data_dimisionX)

model_colorForm <- lmer(transli~relevant_property*dist
             +(1|id)
             +(1|item), data=data_colorForm)

# Check the interaction between relevant_property and dist
step(model_dimisionX) # Sig. Interaction between relevant_property and dist. -> Further Anylysis to explain it needed.
step(model_colorForm) # No sig. Interaction between relevant_property and dist. -> We can look at the summary now and make conclusions.
summary(model_colorForm) 
# 1. Sharp increase color first ordering in color form; 
# 2. Sharp 
```
```{r}
summary(model_dimisionX)
```

For Dimension X data, we further split the data by distribution.
```{r}
emmeans(model_dimisionX, pairwise~relevant_property * dist)
# first blurred - first sharp     -0.5979 0.308 Inf  -1.944  0.3754
emmeans(model_dimisionX, pairwise~dist)
# blurred - sharp   -0.825 0.286 Inf  -2.885  0.0039
```
```{r}
load("./data_preprocessed_huashan.RData")
model_full <- lmer(prefer_first_1st~relevant_property*combination*dist
             +(1|id)
             +(1|item), data=data)
step(model_full) # sig. interaction relevant:dist

model_full <- update(model_full, .~.-relevant_property:combination:dist - relevant_property:combination - combination:dist)
step(model_full)

```
```{r}
data_dimisionX <- subset(data, combination=="dimension_X")
data_colorForm <- subset(data, combination=="color_form")

model_dimensionX <- lmer(prefer_first_1st~relevant_property*dist
             +(1|id)
             +(1|item), data=data_dimisionX)

summary(model_dimensionX) # relevant_propertyfirst:distsharp     5.7283     1.9311 6171.7906   2.966  0.00302 **

step(model_dimensionX) # sig. interaction relevant:dist)

emmeans(model_dimensionX, pairwise~relevant_property * dist) # first blurred - first sharp      -4.850 3.59 Inf  -1.351  0.7564
```

