---
title: "Simulation Analysis — Run 4"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

# Overview

Full simulation run uses `02-simulation-fast.py` (single-process, batched vmap).

**Parameter sweep**

| Parameter | Values | n |
|---|---|---|
| `nobj` | 2, 6, 10, 14, 18, 22, 26, 30 | 8 |
| `speaker` | incremental, global | 2 |
| `semantics` | recursive, static | 2 |
| `color_semvalue` | 0.90, 0.92, 0.94, 0.96, 0.98 | 5 |
| `k` | 0.1, 0.3, 0.5, 0.7, 0.9 | 5 |
| `wf` | 0.1, 0.2, 0.3, 0.5, 0.8 | 5 |
| `sd_spread` | 2.0 (blurred), 7.75 (baseline), 15.0 (sharp) | 3 |

**Fixed:** `alpha = 1`, `bias = 0`, `world_length = 2`, `sample_size = 1000`

Total rows: 8 × 2 × 2 × 5 × 5 × 5 × 3 × 1000 = 12,000,000

**Outcome variables**

- `probs_big_blue` — L1(referent | "big blue") = P(communicative success, size-first order)
- `probs_blue_big` — L1(referent | "blue big") = P(communicative success, color-first order)
- `size_first_advantage` = `probs_big_blue − probs_blue_big` (positive = size-first preferred)

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
rm(list = ls())
library(tidyverse)

# CSP palette from run_2 analysis
CSP_colors <- c(
  "#7581B3", "#99C2C2", "#C65353", "#E2BA78", "#5C7457", "#575463",
  "#B0B7D4", "#66A3A3", "#DB9494", "#D49735", "#9BB096", "#D4D3D9",
  "#414C76", "#993333"
)
scale_colour_discrete <- function(...) scale_colour_manual(..., values = CSP_colors)
scale_fill_discrete   <- function(...) scale_fill_manual(  ..., values = CSP_colors)

theme_set(
  theme_minimal(base_size = 13) +
    theme(
      strip.background = element_rect(fill = "#E8EAF0", colour = NA),
      panel.grid.minor = element_blank()
    )
)
```

# Load & prepare data

```{r load}
data <- read.csv("simulation_full_run.csv")

# Human-readable factor labels
data <- data %>%
  mutate(
    speaker_label = recode(speaker,
      "incremental_speaker" = "Incremental",
      "global_speaker"      = "Global"
    ),
    semantics_label = recode(semantics,
      "recursive" = "Recursive",
      "static"    = "Static"
    ),
    sd_label = factor(
      sd_spread,
      levels = c(2.0, 7.75, 15.0),
      labels = c("Blurred (σ=2)", "Baseline (σ=7.75)", "Sharp (σ=15)")
    ),
    # key derived variable
    size_first_advantage = probs_big_blue - probs_blue_big
  )

cat("Rows:", nrow(data), "\n")
cat("Columns:", paste(names(data), collapse = ", "), "\n")
glimpse(data)
```

# Descriptive overview: marginal means by speaker

```{r speaker-marginal}
speaker_summary <- data %>%
  group_by(speaker_label) %>%
  summarise(
    mean_bb  = mean(probs_big_blue),
    se_bb    = sd(probs_big_blue) / sqrt(n()),
    mean_bB  = mean(probs_blue_big),
    se_bB    = sd(probs_blue_big) / sqrt(n()),
    mean_adv = mean(size_first_advantage),
    se_adv   = sd(size_first_advantage) / sqrt(n()),
    .groups  = "drop"
  )

speaker_summary %>%
  pivot_longer(
    cols      = -speaker_label,
    names_to  = c(".value", "utterance"),
    names_pattern = "(mean|se)_(.*)"
  ) %>%
  filter(utterance %in% c("bb", "bB")) %>%
  mutate(utterance = recode(utterance, bb = "big blue", bB = "blue big")) %>%
  ggplot(aes(x = speaker_label, y = mean, colour = utterance)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = mean - 2 * se, ymax = mean + 2 * se), width = 0.15) +
  labs(title  = "Communicative success by speaker (marginal mean ± 2 SE)",
       x = "Speaker model", y = "P(referent | utterance)", colour = "Utterance") +
  ylim(0, 1)
```

# Speaker × sd_spread

```{r speaker-sd}
data %>%
  group_by(speaker_label, sd_label) %>%
  summarise(
    mean_bb = mean(probs_big_blue),
    se_bb   = sd(probs_big_blue) / sqrt(n()),
    mean_bB = mean(probs_blue_big),
    se_bB   = sd(probs_blue_big) / sqrt(n()),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols      = c(mean_bb, mean_bB, se_bb, se_bB),
    names_to  = c(".value", "utterance"),
    names_pattern = "(mean|se)_(.*)"
  ) %>%
  mutate(utterance = recode(utterance, bb = "big blue", bB = "blue big")) %>%
  ggplot(aes(x = sd_label, y = mean, colour = utterance, group = utterance)) +
  geom_point(size = 3) +
  geom_line() +
  geom_errorbar(aes(ymin = mean - 2 * se, ymax = mean + 2 * se), width = 0.15) +
  facet_wrap(~ speaker_label) +
  labs(title  = "Communicative success by sd_spread and speaker",
       x = "Context sharpness (sd_spread)",
       y = "P(referent | utterance)",
       colour = "Utterance") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

# Size-first advantage

Key question: when does "big blue" beat "blue big"?

```{r advantage-overview}
data %>%
  group_by(speaker_label, sd_label) %>%
  summarise(
    mean_adv = mean(size_first_advantage),
    se_adv   = sd(size_first_advantage) / sqrt(n()),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = sd_label, y = mean_adv, fill = speaker_label)) +
  geom_col(position = "dodge") +
  geom_errorbar(
    aes(ymin = mean_adv - 2 * se_adv, ymax = mean_adv + 2 * se_adv),
    position = position_dodge(0.9), width = 0.2
  ) +
  geom_hline(yintercept = 0, lty = 2, colour = "grey30") +
  labs(title   = "Size-first advantage (probs_big_blue − probs_blue_big)",
       subtitle = "Positive = 'big blue' has higher communicative success",
       x = "Context sharpness", y = "Advantage", fill = "Speaker") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

# Effect of nobj

```{r nobj}
p_nobj <- data %>%
  ggplot(aes(x = nobj)) +
  stat_summary(aes(y = probs_big_blue,  colour = "big blue"),
               fun.data = mean_cl_boot) +
  stat_summary(aes(y = probs_blue_big, colour = "blue big"),
               fun.data = mean_cl_boot) +
  facet_grid(sd_label ~ speaker_label) +
  labs(title  = "Communicative success by nobj",
       x      = "N objects in context",
       y      = "P(referent | utterance)",
       colour = "Utterance") +
  scale_x_continuous(breaks = unique(data$nobj))

p_nobj
ggsave("sim_nobj.png", p_nobj, width = 9, height = 7, dpi = 300)
```

# Effect of color_semvalue

```{r csv}
p_csv <- data %>%
  group_by(speaker_label, sd_label, color_semvalue) %>%
  summarise(
    mean_bb = mean(probs_big_blue),
    se_bb   = sd(probs_big_blue) / sqrt(n()),
    mean_bB = mean(probs_blue_big),
    se_bB   = sd(probs_blue_big) / sqrt(n()),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols      = c(mean_bb, mean_bB, se_bb, se_bB),
    names_to  = c(".value", "utterance"),
    names_pattern = "(mean|se)_(.*)"
  ) %>%
  mutate(utterance = recode(utterance, bb = "big blue", bB = "blue big")) %>%
  ggplot(aes(x = color_semvalue, y = mean, colour = utterance)) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymin = mean - 2 * se, ymax = mean + 2 * se), width = 0.005) +
  facet_grid(sd_label ~ speaker_label) +
  labs(title  = "Communicative success by color_semvalue",
       x = "Color semantic value", y = "P(referent | utterance)",
       colour = "Utterance")

p_csv
ggsave("sim_colorsemval.png", p_csv, width = 9, height = 8, dpi = 300)
```

# Effect of k (threshold parameter)

```{r k}
p_k <- data %>%
  group_by(speaker_label, sd_label, k) %>%
  summarise(
    mean_bb = mean(probs_big_blue),
    se_bb   = sd(probs_big_blue) / sqrt(n()),
    mean_bB = mean(probs_blue_big),
    se_bB   = sd(probs_blue_big) / sqrt(n()),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols      = c(mean_bb, mean_bB, se_bb, se_bB),
    names_to  = c(".value", "utterance"),
    names_pattern = "(mean|se)_(.*)"
  ) %>%
  mutate(utterance = recode(utterance, bb = "big blue", bB = "blue big")) %>%
  ggplot(aes(x = k, y = mean, colour = utterance)) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymin = mean - 2 * se, ymax = mean + 2 * se), width = 0.02) +
  facet_grid(sd_label ~ speaker_label) +
  labs(title  = "Communicative success by k (threshold strictness)",
       x = "k", y = "P(referent | utterance)", colour = "Utterance")

p_k
ggsave("sim_k.png", p_k, width = 9, height = 8, dpi = 300)
```

# Effect of wf (perceptual blur)

`wf` controls the steepness of the size CDF: low wf = sharp/certain size semantics; high wf = blurred/uncertain size semantics. Lower wf should make "big" a more reliable first word for the incremental speaker.

```{r wf}
p_wf <- data %>%
  group_by(speaker_label, sd_label, wf) %>%
  summarise(
    mean_bb = mean(probs_big_blue),
    se_bb   = sd(probs_big_blue) / sqrt(n()),
    mean_bB = mean(probs_blue_big),
    se_bB   = sd(probs_blue_big) / sqrt(n()),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols      = c(mean_bb, mean_bB, se_bb, se_bB),
    names_to  = c(".value", "utterance"),
    names_pattern = "(mean|se)_(.*)"
  ) %>%
  mutate(utterance = recode(utterance, bb = "big blue", bB = "blue big")) %>%
  ggplot(aes(x = wf, y = mean, colour = utterance)) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymin = mean - 2 * se, ymax = mean + 2 * se), width = 0.02) +
  facet_grid(sd_label ~ speaker_label) +
  labs(title  = "Communicative success by wf (perceptual blur)",
       subtitle = "Lower wf = sharper size semantics",
       x = "wf", y = "P(referent | utterance)", colour = "Utterance")

p_wf
ggsave("sim_wf.png", p_wf, width = 9, height = 8, dpi = 300)
```

# Size-first advantage by wf

```{r adv-wf}
p_adv_wf <- data %>%
  group_by(speaker_label, wf, sd_label) %>%
  summarise(
    mean_adv = mean(size_first_advantage),
    se_adv   = sd(size_first_advantage) / sqrt(n()),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = wf, y = mean_adv, colour = sd_label, group = sd_label)) +
  geom_point(size = 2.5) +
  geom_line() +
  geom_ribbon(aes(ymin = mean_adv - 2 * se_adv,
                  ymax = mean_adv + 2 * se_adv,
                  fill = sd_label), alpha = 0.15, colour = NA) +
  geom_hline(yintercept = 0, lty = 2, colour = "grey30") +
  facet_wrap(~ speaker_label) +
  labs(title   = "Size-first advantage by wf and sharpness",
       subtitle = "Crosses zero = incremental flips to prefer big-blue",
       x = "wf", y = "Advantage (big blue − blue big)",
       colour = "Context", fill = "Context")

p_adv_wf
ggsave("sim_advantage_wf_sd.png", p_adv_wf, width = 9, height = 5, dpi = 300)
```

# Size-first advantage: k × sd_spread heatmap

```{r heatmap}
p_heat <- data %>%
  group_by(speaker_label, k, sd_label) %>%
  summarise(mean_adv = mean(size_first_advantage), .groups = "drop") %>%
  ggplot(aes(x = factor(k), y = sd_label, fill = mean_adv)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = round(mean_adv, 3)), size = 3) +
  scale_fill_gradient2(low = CSP_colors[3], mid = "white", high = CSP_colors[1],
                       midpoint = 0, name = "Advantage") +
  facet_wrap(~ speaker_label) +
  labs(title = "Mean size-first advantage: k × sd_spread",
       x = "k", y = "Context sharpness")

p_heat
ggsave("sim_heatmap_k_sd.png", p_heat, width = 10, height = 4, dpi = 300)
```

# Size-first advantage: color_semvalue × sd_spread heatmap

```{r heatmap2}
p_heat2 <- data %>%
  group_by(speaker_label, color_semvalue, sd_label) %>%
  summarise(mean_adv = mean(size_first_advantage), .groups = "drop") %>%
  ggplot(aes(x = factor(color_semvalue), y = sd_label, fill = mean_adv)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = round(mean_adv, 3)), size = 3) +
  scale_fill_gradient2(low = CSP_colors[3], mid = "white", high = CSP_colors[1],
                       midpoint = 0, name = "Advantage") +
  facet_wrap(~ speaker_label) +
  labs(title = "Mean size-first advantage: color_semvalue × sd_spread",
       x = "color_semvalue", y = "Context sharpness")

p_heat2
ggsave("sim_heatmap_csv_sd.png", p_heat2, width = 10, height = 4, dpi = 300)
```

# nobj × sd_spread interaction (size-first advantage)

```{r adv-nobj-sd}
p_inter <- data %>%
  group_by(speaker_label, semantics_label, nobj, sd_label) %>%
  summarise(
    mean_adv = mean(size_first_advantage),
    se_adv   = sd(size_first_advantage) / sqrt(n()),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = nobj, y = mean_adv, colour = sd_label, group = sd_label)) +
  geom_point(size = 2.5) +
  geom_line() +
  geom_ribbon(aes(ymin = mean_adv - 2 * se_adv,
                  ymax = mean_adv + 2 * se_adv,
                  fill = sd_label), alpha = 0.15, colour = NA) +
  geom_hline(yintercept = 0, lty = 2, colour = "grey30") +
  facet_grid(semantics_label ~ speaker_label) +
  scale_x_continuous(breaks = unique(data$nobj)) +
  labs(title   = "Size-first advantage by nobj and sharpness",
       x = "N objects", y = "Advantage (big blue − blue big)",
       colour = "Context", fill = "Context")

p_inter
ggsave("sim_advantage_nobj_sd.png", p_inter, width = 9, height = 7, dpi = 300)
```

# Descriptive summary table

```{r summary-table}
data %>%
  group_by(speaker_label, sd_label) %>%
  summarise(
    n            = n(),
    mean_bb      = round(mean(probs_big_blue),  4),
    sd_bb        = round(sd(probs_big_blue),    4),
    mean_bB      = round(mean(probs_blue_big),  4),
    sd_bB        = round(sd(probs_blue_big),    4),
    mean_adv     = round(mean(size_first_advantage), 4),
    sd_adv       = round(sd(size_first_advantage),   4),
    .groups = "drop"
  ) %>%
  knitr::kable(col.names = c(
    "Speaker", "Context", "N",
    "Mean P(bb)", "SD P(bb)",
    "Mean P(bB)", "SD P(bB)",
    "Mean adv.", "SD adv."
  ))
```
