---
title: "Analysis"
author: "Your Name"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---
# Simulation with Random States

**Single object $O$:**

A single object is a tuple of $<s, c, f>$, where:

- The size $s$ is sampled from a size distribution. There are three types of distributions: Normal, Left-Skewed, and Right-Skewed. All distributions are truncated within the range $[1, 30]$.
  - Normal: $s \sim N(\mu = 15, \sigma = 7.5)$
  - Left-Skewed: $s \sim N(\mu = 22.5, \sigma = 7.5)$
  - Right-Skewed: $s \sim N(\mu = 7.5, \sigma = 7.5)$
- Both color $c$ and form $f$ are sampled from a Bernoulli distribution: 

  $c$ or $f$ $\sim Bern(p = 0.5)$

**Single context $C$:**

A single context consists of $nobj$ objects, which serve as the input for the model to assess. It is analogous to a single trial in a behavioral experiment.

**A referent $r$** is defined as follows:

$$
<s_r, c_r, f_r>: s_r = \arg\max s \in C; c_r = f_r = 1.
$$

**Modification on the core model:**

1. The method for computing the threshold is now sample-based.
2. We added a pragmatic listener model.

**A communicative success** is defined as follows:

Variant 1:
$$
1 \iff L_1(r|u_1, C) > L_1(r|u_2, C), r \in C
$$
where $u_1$ is "big blue" and $u_2$ is "blue big". $L_1$, the pragmatic listener, is defined as:
$$
L_1(r|u,C) \propto S_1(u|r,C) \cdot P(r)
$$

This is equivalent to variant 2:
$$
1 \iff S_1(u_1|r, C) > S_1(u_2|r, C), r \in C
$$

Variant 3:
$$
1 \iff \operatorname*{argmax}_r L_1(r|u_1,C)
$$

**Single simulation run:**

With one unique parameter setting, the sample size is 10,000 (1e4). This corresponds to the number of all single contexts (trials). The parameter values vary across different simulation runs and are described as follows:

- nobj - [2,6,...,18], step = 4. Number of objects in a single context, 5 values in total.
- speaker - incremental speaker or global speaker. List of speaker models, 2 values in total.
- color semval - [0.90,0.92,...,0.99], step = 0.02. Semantic values for color adjectives, 6 values in total.
- k - [0.2,0.4,...,0.8], step = 0.2. Percentage for determining the threshold for size semantics, four values in total.
- wf - [0.2,0.4,...,1], step = 0.2. Parameter for perceptual blur, five values in total.
- size distribution - normal, left-skewed, and right-skewed. Size distribution for sampling size for a single object, 3 values in total.

This results in a total of 1800 iterations, and 18 million samples.

**Initial results:**

1. Incremental speaker generally leads to less communicative success with size-first ordering.
2. The intention of manipulating the size distribution is to manipulate the scale of communicative efficiency of size adjectives, from low to high: right-skewed, normal, left-skewed. This manipulation seems not very effective. The intended communicative success is the deviation of the real size value from the threshold. The real value is max. The threshold is a function of max, min, and k. Given a fixed k, to achieve a low threshold, we must maximize max and the difference between max and min. The current manipulation does not effectively achieve this. Currently, left-skewed is the lowest, and normal and right-skewed are almost equal.
3. The higher the wf, k, and color semval are, the lower the communicative success.

## Introduction

This is an analysis of the data.


Load the data.
```{r}
rm(list=ls())
filename <- "simulation_full_run_2.csv"
data <- read.csv(filename)
```


```{r}
data$nobj <- data$nobj / 10000
```
Set up the theme for plot.
```{r}
library(tidyverse)
library(aida)   # custom helpers: https://github.com/michael-franke/aida-package
theme_set(theme_aida())

##################################################
## CSP-colors
##################################################
CSP_colors = c(
  "#7581B3", "#99C2C2", "#C65353", "#E2BA78", "#5C7457", "#575463",
  "#B0B7D4", "#66A3A3", "#DB9494", "#D49735", "#9BB096", "#D4D3D9",
  "#414C76", "#993333"
  )
# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = CSP_colors)
}
scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = CSP_colors)
}
CSP_color_names = c("glaucous", "opal", "shimmer", "crayola", "fern", "independence",
           "glaucous light 2", "opal dark 2", "shimmer light 2", "crayola dark 2", "fern light 3", "independence light 4",
           "glaucous dark 3", "shimmer dark 2"
           )
```
Display the difference between two speakers and two ordering.

```{r}


p <- data %>% 
  ggplot(aes(x=speaker, utt.probs)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_big_blue), color = CSP_colors[1], show.legend=TRUE) +
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_blue_big), color = CSP_colors[3], show.legend=TRUE) +
  scale_y_continuous(sec.axis = sec_axis(~ .)) +
  scale_color_manual(
                     name = "Probability of recovering r") +
  theme_aida()

ggsave("sim_speaker_plot.png", p, width = 8, height = 6, dpi = 300)
```
Display the error bar for big blue.
```{r}
p <- data %>% 
  ggplot(aes(x=speaker, utt.probs)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_big_blue), color = CSP_colors[1]) +
  scale_y_continuous(sec.axis = sec_axis(~ .)) +
  theme_aida()

ggsave("sim_speaker_uncertainty_plot.png", p, width = 8, height = 6, dpi = 300)
```

```{r}
p <- data %>% 
  ggplot(aes(x=nobj, utt.probs)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_big_blue), color = CSP_colors[1]) +
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_blue_big), color = CSP_colors[3]) +
  scale_y_continuous(sec.axis = sec_axis(~ .)) +
  facet_wrap(~speaker) +
  theme_aida()

ggsave("sim_nobj_plot.png", p, width = 8, height = 6, dpi = 300)
```

```{r}
p <- data %>% 
  ggplot(aes(x=color_semvalue, utt.probs)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_big_blue), color = CSP_colors[1]) +
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_blue_big), color = CSP_colors[3]) +
  scale_y_continuous(sec.axis = sec_axis(~ .)) +
  facet_grid(~speaker) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_aida()

ggsave("sim_colorsemval_plot.png", p, width = 8, height = 6, dpi = 300)
```

```{r}
p <- data %>% 
  ggplot(aes(x=wf, utt.probs)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_big_blue), color = CSP_colors[1]) +
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_blue_big), color = CSP_colors[3]) +
  scale_y_continuous(sec.axis = sec_axis(~ .)) +
  facet_grid(~speaker) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_aida()

ggsave("sim_wf_plot.png", p, width = 8, height = 6, dpi = 300)
```

```{r}
p <- data %>% 
  ggplot(aes(x=k, utt.probs)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_big_blue), color = CSP_colors[1]) +
  stat_summary(fun.data = "mean_cl_boot", aes(y = probs_blue_big), color = CSP_colors[3]) +
  scale_y_continuous(sec.axis = sec_axis(~ .)) +
  facet_grid(~speaker) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme_aida()

ggsave("sim_k_plot.png", p, width = 8, height = 6, dpi = 300)
```

```{r}
# descriptive statistics
data %>% 
  group_by(speaker, nobj) %>% 
  summarise(
    mean = mean(probs_big_blue),
    sd = sd(probs_big_blue),
    n = n()
  )
```
## Legacy: First test run with coded comminicative sucess
Load the data.
```{r}
filename <- "simulation_test_run_fewer_parameters.csv"
data <- read.csv(filename)

```
Set up the theme for plot.
```{r}
library(tidyverse)
library(aida)   # custom helpers: https://github.com/michael-franke/aida-package
theme_set(theme_aida())

##################################################
## CSP-colors
##################################################
CSP_colors = c(
  "#7581B3", "#99C2C2", "#C65353", "#E2BA78", "#5C7457", "#575463",
  "#B0B7D4", "#66A3A3", "#DB9494", "#D49735", "#9BB096", "#D4D3D9",
  "#414C76", "#993333"
  )
# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = CSP_colors)
}
scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = CSP_colors)
}
CSP_color_names = c("glaucous", "opal", "shimmer", "crayola", "fern", "independence",
           "glaucous light 2", "opal dark 2", "shimmer light 2", "crayola dark 2", "fern light 3", "independence light 4",
           "glaucous dark 3", "shimmer dark 2"
           )
```

## Plot the propotion of sucess again color_semvalue & speaker

```{r}
data %>% 
  ggplot(aes(x=speaker, y=proportion_success)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot") +
  theme_aida()
```
```{r}
data %>%
  ggplot(aes(x=size_distribution, y=proportion_success)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot") +
  facet_wrap(~speaker) + 
  theme_aida()
```

To gain a basic understanding of the data and the impact of parameter values.
```{r}
data %>% filter(size_distribution == "normal") %>%
  ggplot(aes(x=color_semvalue, y=proportion_success)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot") +
  facet_wrap(~speaker) + 
  theme_aida()
```
## Plot the propotion of sucess again nobj & speaker
```{r}
data %>% filter(size_distribution == "normal") %>%
  ggplot(aes(x=nobj, y=proportion_success)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot") +
  facet_wrap(~speaker) + 
  theme_aida()
```
```{r}
p <- data %>% filter(size_distribution == "normal") %>%
  ggplot(aes(x=wf, y=proportion_success)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot") +
  facet_wrap(~speaker) + 
  ylab("probs.subj.first.preferred") +
  theme_aida()

ggsave("sim_wf_plot.png", p, width = 8, height = 6, dpi = 300)
p
```
```{r}
p <- data %>% filter(size_distribution == "normal") %>%
  ggplot(aes(x=k, y=proportion_success)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot") +
  facet_wrap(~speaker) + 
  ylab("probs.subj.first.preferred") +
  theme_aida()

ggsave("sim_k_plot.png", p, width = 8, height = 6, dpi = 300)
p

```

```{r}
library(ggplot2)
#ggplot(data, aes(x=color_semvalue, y=proportion_success)) + geom_smooth()
data %>% filter(wf == 0.6 & k == 0.6 & size_distribution == "normal" & nobj == 10) %>%
  ggplot(aes(x=nobj, y=proportion_success)) + 
  #geom_point(stat="identity", position=position_dodge()) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  #facet_wrap(~speaker) + 
  facet_grid(speaker~color_semvalue) +
  theme_aida()

```

